### 2020-01-26 latex/sections/introduction.tex
updating skeleton

In the year $2017$ the approximate amount of adults with diabetes was 425-million people \cite{Intro_FactsandFigures}, this number is 
projected to rise to $629$-million people by the year $2045$. The rate of people that are diagnosed with diabetes 
is 1 in every 2. Secondly of all people with diabetes 1 in every 3 develop diabetic retinopaty (DR)\cite{Intro_DiabetesRate}.

\textit{Diabetic retinopathy} is a symptom of both type 1 and type 2 diabetes and is now the leading cause of blindness amongst adults today. DR is 
diagnosed by two major technologies, firstly there is the fundus imaging method, where a photograph is taken of the fundus (back) 
of the patients eye, an example of this can be seen in Figure \ref{fig:fundus-image1}. Secondly ophthalmologists examine 
Optical Choerence Tomography (OCT) scans of human retina, an example of this can be seen in Figure \ref{fig:oct-retina1}. 
The work in this paper does not focus on the latter. The diagnosis of DR is costly and as the statistic indicate 
the amount of work is going to increase.

\begin{figure}[h]
  \centering
  \includegraphics[width=.3\linewidth]{../graphics/fundus_image1.jpg}
  \caption{Example of Fundus image}
  \label{fig:fundus-image1}
\end{figure}
\begin{figure}[h]
  \centering
  \includegraphics[width=.5\linewidth]{../graphics/oct_fundus.png}
  \caption{Example of OCT on Retina}
  \label{fig:oct-retina1}
\end{figure}

Patients that have DR are grouped into four different stages\cite{Intro_FourStages}:

1. \textit{Mild Nonproliferative Retinopathy} - This is the earliest stage of diabetic retinopathy, and it’s characterized by balloon-like swelling in the retina’s blood vessels. These are called microaneurysms, and these vessels can leak into the eye.

2. \textit{Moderate Nonproliferative Retinopathy} - During this stage, the blood vessels nourishing the retina swell and may even become blocked. This can contribute to diabetic macular edema (DME) which is a build-up of fluid in the macula region of the retina.

3. \textit{Severe Nonproliferative Retinopathy} -  At this stage, an increasing number of blood vessels nourishing the eye have become blocked. As a result, the retina is signaled to grow new blood vessels.

4. \textit{Proliferative Diabetic Retinopathy} - This is the final stage of diabetic retinopathy. New blood vessels proliferate, growing inside the retina and into the vitreous gel, which is the fluid that fills the eye. Because these blood vessels are delicate, they may begin to leak and bleed. As a result, scar tissue may form, causing retinal detachment, the pulling away of the retina from underlying tissue. Retinal detachment may cause spotty vision, flashes of lights, or severe vision loss.

Generally without intervention patients advance from one stage to the next eventually leading 
to permanent vision impairment, however with intervention by an ophthalmologist this advancement 
can be prevented \cite{Intro_StageAdvancement}.

This paper focuses on applying machine learning methods to patient information and Fundus images of those 
patients to assess the likelyhood (risk) of those patients DR developing into SDR.

In this paper we will apply machine learning methods to patients information and fundus images of patients 
that has been collected a period of time to examine a patiens likelyhood to advance from one stage 
of DR to the next, attempting to predict the timeline the patient has to seek medical help.

\section{Machine learning methods in Ophthalmology}

Applying machine learning to fundus images and patient information has been done before very successfully 
researchers\cite{Intro_googleDeepmind}, this work generally focuses on applying image recognition on 
the images to identify which stage of DR the eye has reached.

\section{Background}
\ifdraft{Provide background about the subject matter (e.g. How was morse code
developed?  How is it used today?). 
This is a place where there are usually many citations.
It is suspicious when there is not. 
Include the purpose of the different equipment and your design intent. 
Include references to relevant scientific/technical work and books.}

Applying statistics to medical data in order to better understand a patients evolution from DR to SDR 
has been done previously by a company called RetinaRisk, a significant contributor to this paper, but 
RetinaRisk's solution does not apply machine learning but a specialists field knowledge and a handcrafted 
formula. RetinaRisks formula has been used to significantly reduce costs to health organizations as 
when patients following their recommendations for doctor checkups come in less often but at time that 
are more critical. im just writing something so that it looks like I'm writing something

>>>>

In the world of deep learning is exciting, we have models that are able to examine
images and very reliably be able to identify it's content. Deep learning models have
beaten Opthalmologists in identifing diabetic retinopathy, they've identified
cancer cells where others have not. Deep learning models have even predicted
the likelyhood a person will die in the following year. These models have done
these things extremely accurately, cheap, and fast. These successes allow us to
dream of a future where all things are eventually possible with the help of
technology.

The field of examining deep learning models to gain a richer understanding
in how they work, and what makes them so much better than humans at various tasks
is still new. This is the field of Explainable Artificial Intelligence (XAI).
If we wish to continue using artificial intelligence (AI) and machine learning (ML),
to improve our lives we must examine how they work, this is not only to improve
ourselves but these explanations are a legal requirement now in many areas, namely, legal, and
medical.

This thesis takes a look at how we can examine an artificial neural network (ANN)
to understand which higher level concepts (HLC) it deems important for a given state
within games. These games can be simple like Tic-Tac-Toe or Breakthrough and extremely
complicated like Chess or Go. If an ANN values a HLC greatly which we as humans don't find
very important, and yet the ANN is able to defeat us there is reason to think
that HLC should be of more importance to us.

The chess engine DeepBlue was the first chess engine to win a chess match agains a
chess grand master. It's power in the world of chess was a marvel. DeepBlue used tree-
search to evaluate it's position and possible moves, doing an immense amount of calcluations
in order to arrive at the best move it could given the time constraints. The tree search
approach is dependent on knowing which board state is good and what board state is bad,
This can be easy when the boardstate is such that you're guaranteed to have been checkmated
in one move, and it's incredibly hard when you boardstate is such that both you and your
opponent have 16 pieces and are only a handfull of moves in. Generally the boardstate
is a function of the pieces you have, how safe they are, and how opportunistic you are to
capture your opponents pieces. These values are not obvious and require a domain expert
to describe the situations and give the values for these cases.

The process of evaluating chess positions did not change for many years. With the revival
of deep neural networks this changed. A neural network is a network of connected nodes (neurons)
that each how a specific weight. These networks are fed input which is passed through the
network and it is updated by the weights of the neurons, the input is morphed into
a more comprehensible value and these values are then used to generate some output.
The output can be a binary classification for instance input=Image output=\{cat, dog\}.
Processing the pixels and altering them to numerical values to find the output cat or dog
is not intuitive and understanding of the values of the neurons in the network
is not considered fruitful.

This approach of
chess engines is still popular today but not the reigning champion anymore. In todays
best chess engines deep neural networks are used to

Deep neural network machine learning models have an amazing impact on modern society today,
we use them to know which items to purchase, to know which people we should love, and we
use them to drive us to work. These tasks given to deep learning models are all niceties
for us to enjoy. However, as our dependance on them increases and their astounding success
in any and all tasks continues, we will want to utilize them for all things. This leads
to a utopia where decision making is removed from the hands of humans. The question we
must ask ourselves is when do we want to know why a decision was taken, is it only when
the decision taken is not preferrable to us. For instance, if a deep neural network model
used by your lending authority decides you are not worty of a loan. These questions should
have answers but as it stands today no such answer is possible.

We can use methods like saliency maps to map a deep learning model emphasis on pixels to
hopefully understand what portions the model is focusing on. This method however, tells us
nothing about what

results that are nothing but amazing.

### 2020-01-26 latex/sections/managerial.tex
updating skeleton

%% Dedication is optional, comment out if it is absent
\begin{dedications}
  I dedicate this to my spouse/child/pet/power animal.
\end{dedications}

@@

%% The list of abbreviations is an example of a special list
%% Other lists may be added, such as lists of algorithms, symbols, theorems, etc.
%% IN CS PhD, this is sometimes centered.
\coverchapter{List of Abbreviations}%%RUM: Not mentioned
\begin{tabular}{ll}
MSc &Masters of Science\\
DR & Diabetic Retinopathy\\
ML & Machine Learning\\
AI & Artificial Intelligence\\
\end{tabular}

% \coverchapter{List of Symbols}%%RUM: Not mentioned
% \begin{tabular}{lll}
% Symbol &Description &Value/Units\\
% $E$ &Energy &\si{\joule}\\
% $m$ &Mass &\si{gram}\\
% $c$ &Speed of Light &\SI{2.99E8}{\meter\per\second\square}\\
% \end{tabular}

>>>>

%% Dedication is optional, comment out if it is absent
\begin{dedications}
  I dedicate this thesis to my family who while not understanding most of what I do
  always support me. The friends that still want to talk to me even though my
  schedule has rendered me unmeetupwithable. Lastly but certainly not least,
  Hulda Lilja who always easead my mind during my spurts of imposter syndrome, and fear of
  failure.

  Don't Panic.
\end{dedications}

@@

%% The list of abbreviations is an example of a special list
%% Other lists may be added, such as lists of algorithms, symbols, theorems, etc.
%% IN CS PhD, this is sometimes centered.
\coverchapter{List of Abbreviations}%%RUM: Not mentioned
\begin{tabular}{ll}
MSc &Masters of Science\\
ML & Machine Learning\\
AI & Artificial Intelligence\\
\end{tabular}
